Equivalence Hypergraphs: E-Graphs for Monoidal Theories

### 2024-01-05 E-Hypergraphs-LICS24.tex
fix figures, section 3.1 blurb.

\begin{figure*}\label{fig:e-graph-example}
\[
	\hspace{2cm}
    \scalebox{0.65}{
    \tikzfig{categorical-semantics/egraph-translation}
    }
\]
\caption{Example translation of acyclic e-graphs into string diagrams for semilattice enriched Cartesian categories. }
\end{figure*}

@@

\begin{figure}\label{fig:egraph-strings}
\[
	\scalebox{0.65}{
	\tikzfig{categorical-semantics/egraph-strings}
	}
\]
\caption{String diagrams for semilattice enriched symmetric monoidal categories.}
\end{figure}

@@

Our use of string diagrams is also central to the intuition behind our work.  
DAN: INFORMAL REASONING DEVICES DOESN'T SOUND GOOD. WHAT DO YOU MEAN? 
The string diagrams we work with are left essentially as informal reasoning devices, however there exists a breadth of work on similar diagrammatic formalisms for similar classes of categories.  We mention here work on hierarchical string diagrams \cite{ghica_hierarchical_2023},  functorial boxes \cite{mellies_functorial_2006},  proof nets for compact closed categories with biproducts \cite{duncan_generalised_2009}, and recent work on tape diagrams for rig categories with biproducts \cite{bonchi_tape_nodate}.

DAN: THIS NEEDS TO BE CAREFULLY EXPANDED TO NOT OFFEND POTENTIAL REVIEWERS.

\section{Categorical Semantics of E-Graphs}
CHRIS: INTRO TO SECTION: CLARIFY THAT WE GIVE E-GRAPHS AS CARTESIAN CASE,  BUT ALSO GENERALIZE TO MONOIDAL CASE. 
CHRIS: CITE HASEGAWA - TRACE MONOIDAL CATEGORIES, AND PERDIX. 
DAN: I SUGGEST TO CITE HASEGAWA IN FURTHER WORK, FOR GENERALISING THE THEORY TO TRACED MC.

@@

A categorical presentation of a monoidal signature is given by a freely generated products and permutations category.  
\begin{definition}[Products and permutations category]
A \textit{products and permutations category (PROP)} is a strict symmetric monoidal category where every object is the n-fold tensor product of some distinguished object $A$. In this case, we will denote the n-fold tensor product of the distinguished object by simply $n$, so that $n \otimes m = n+m$. A \textit{PROP functor} is a strict SMC-functor which is additionally identity-on-objects. 
We denote the free \textit{products and permutations category} over a monoidal signature $\Sigma$ by $\textsf{PROP}(\Sigma)$.  
\end{definition}
The free category $\textsf{PROP}(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $f \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$, quotiented by the axioms of a symmetric monoidal category.

An equation associated with a monoidal signature is a pair of parallel morphisms of the form above, leading to the following definition of a symmetric monoidal theory. 
\begin{definition}[Symmetric Monoidal Theory]
A \textit{presentation of a symmetric monoidal theory} $(\Sigma, \mathcal{E})$ is given by a pair of a monoidal signature $\Sigma$ and a set $\mathcal{E}$ of pairs of parallel morphisms $\phi,\psi: n \to m$ of $\textsf{PROP}(\Sigma)$.
A \textit{symmetric monoidal theory} $\textsf{SMT}(\Sigma,\mathcal{E})$ generated by a presentation $(\Sigma, \mathcal{E})$ is given by $\textsf{PROP}(\Sigma)$ quotiented by the least congruence including $\mathcal{E}$.
\end{definition}

It will often be convenient to give presentations of SMTs in terms of string diagrams.
\begin{example}
The SMT of \textit{commutative comonoids} is given by the following generators, ${\Delta, !}$, depicted in string diagrammatic form:
\[
	\scalebox{0.6}{
  	 \tikzfig{categorical-semantics/Cartesian-equipment}
	}
\]
together with the following associativity, commutativity and unitality equations, again given in terms of string diagrams. 
\[
	\scalebox{0.6}{
	\tikzfig{categorical-semantics/Cartesian-theory}	
	}
\]
The Cartesian SMT is given by a set of generators $\Sigma_C$, together with the generators and equations of the SMT of commutative monoids, and the following additional naturality equations, for every $c \in \Sigma_C$
\[
	\scalebox{0.6}{
	\tikzfig{categorical-semantics/Cartesian-naturality}
	}
\]
By Fox's Theorem~cite{cite},  $\textsf{SMT}(\Sigma_C, \mathcal{E}_C)$ is equivalent to the free Cartesian category over $\Sigma_C$. 
\end{example}

@@

We will use enrichment to axiomatize the e-box structure as the ``join" of two morphisms $f,g: A \to B$ as $f + g: A \to B$.\footnote{Note, in particular,  that semilattices do not require a unit for $+$. This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must also be a categorical coproduct---a degeneracy we wish to avoid. DAN: REFERENCE OR PROOF?} More precisely our hom-sets will be given the structure of a semilattice.  
Generally,  we may define a category \textit{enriched} in any other monoidal category $M$ (\textit{e.g.},  \textbf{SLatt}).  In this case,  hom-sets are generalized to hom-objects of $M$ (\textit{e.g.},  a semilattice) and composition 
\[
	\circ: Hom(B,C) \otimes Hom(A,B) \to Hom(A,C)
\]
is defined as a $M$-morphism.  In other words,  composition must respect the additional structure on hom-sets.  
In our case,  the category we wish to enrich is also monoidal,  so we additionally ask that the monoidal structure also respects this structure.  Technically, this amounts to asking that the tensor product is an enriched functor. 
We omit further details of the general case of enrichment here, and instead work with the following more concrete defintion, where $M = \textbf{SLatt}$.

\begin{definition}[Semilattice-enriched PROP]
A \textit{semilattice enriched strict SMC (PROP)}  $\mathbb{C}$ is given by an SMC (PROP) $(\mathbb{C}, \otimes, 1,+)$ where every hom-set additionally has the structure of a semilattice $(\mathbb{C}(A,B), +)$ which respects the composition and tensor product in the following ways,  for all appropriately typed morphisms. 
\begin{align*}
f ; (g+h) &= f;g + f;h &
(f+g) ; h &= f;h + g;h \\
f \otimes (g+h) &= f \otimes g + f \otimes h & 
(f+g) \otimes h &= f \otimes h + g \otimes h,
%\item A set $ob(\mathbb{C})$ of objects;
%\item For every pair $A,B \in ob(\mathbb{C})$, a hom-semigroup $Hom(A,B)$;
%\item An element $\textsf{id}_{A,B} \in Hom(A,B)$; 
\end{align*}
Note,  we take $(;\!)$ and $\otimes$ to bind more tightly than $+$.
A \textit{semilattice enriched strict SMC (PROP) functor} $F: (\mathbb{C}, \otimes, 1,+) \to (\mathbb{D}, \otimes, 1,+)$ is given by a strict SMC (PROP) functor on the underlying SMCs (PROPs) which additionally satisfies $F(f+g) = F(f)+F(g)$.  We denote the \textit{freely enriched PROP (SMT)} over a monoidal signature $\Sigma$ as $\textsf{PROP}^+(\Sigma)$ ($\textsf{SMT}^+(\Sigma, \mathcal{E})$,  respectively). 
\end{definition}

The free category $\textsf{PROP}^+(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $f \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$, as well as $f+g$, quotiented by the axioms of an enriched symmetric monoidal category.

@@

In this section,  we recall the fundamental definitions and results of string diagram rewrite theory for symmetric monoidal categories,  building to the recapitulation of the equivalence between hypergraphs of a particular form and terms in $\textsf{PROP}(\Sigma)$ established in~\cite{Frobenius, Frobenius2},  and how one can implement string diagram rewriting using DPO rewriting on hypergraphs.   
CHRIS: INTRO HYPERGRAPHS WITH INTERFACES,  RESTRICTIONS TO SMCS.  Q: INTRO DPO REWRITING HERE TOO? NOTE,   REMARK THAT WE CANNOT DEAL WITH SIGNATURES WITH $f:0 \to 0$ morphisms!

@@

Hypergraphs generalize graphs by allowing edges to have multiple sources and targets. 
\begin{definition}[Category of hypergraphs]\label{def:hypergraph}
A \emph{hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l)$,  where $V$ is a finite set of nodes, $E$ is a finite set of edges, $s : E \to V^{*}$ is a source function, $t : E \to V^{*}$ is a target function,  and $l : E \to \Sigma$ is a labelling function that assigns each edge a generator from monoidal signature $\Sigma$.  The labelling function must respect the typing of the generator: for all edges $e$,  $|s(e)| = m$ and $|t(e)| = n$,  where $l(e) : m \to n$. 
A \emph{hypergraph homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ is given by a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that the following hold. 
\begin{enumerate}
    \item $(s_{\mathcal{F}}\, ;\phi_V^*)(e) = (\phi_E\, ;s_{\mathcal{G}})(e)$
    \item $(t_{\mathcal{F}}\, ;\phi_V^*)(e) = (\phi_E\, ;t_{\mathcal{G}})(e)$
    \item $l_{\mathcal{F}}(e) = \phi_E;l_{\mathcal{G}}(e)$
\end{enumerate}
We denote by $\catname{Hyp(\Sigma)}$ the category of hypergraphs over $\Sigma$.
\end{definition}
Note that $\catname{Hyp(\Sigma)}$ has all finite colimits and, in particular,  the coproduct is given by the disjoint union of hypergraphs.
The initial object is given by the empty hypergraph.  
CHRIS: DESCRIPTION OF PUSHOUT HERE?

When we interpret string diagrams as hypergraphs,  we will interpret morphisms as edges,  and wires by vertices.  However,  we will additionally need a way to specify which vertices are to be considered as \textit{input} and \textit{output}. 
For this purpose,  will use the following construction to equip our hypergraphs with \textit{interfaces}. 
\begin{definition}[Symmetric monoidal category of cospans]
Let $\mathbb{C}$ be a category with all finite colimits.  A cospan from $X$ to $Y$ is a pair of arrows $X \xrightarrow{} A \xleftarrow{} Y$  in $\mathbb{C}$.  This \textit{category of cospans of $\mathbb{C}$},  denoted $\catname{Csp(\mathbb{C})}$,  has as objects those of $\mathbb{C}$ and as arrows isomorphism classes of cospans.
%Two cospans $X \xrightarrow{} A \xleftarrow{} Y$ and $X \xrightarrow{} B \xleftarrow{} Y$ are isomorphic if the following diagram commutes and $\alpha$ is iso.
%\[
%\begin{tikzcd}
%                                               & A \arrow[dd, "\alpha"] &                                                \\
%X \arrow[ru, bend left] \arrow[rd, bend right] &                        & Y \arrow[ld, bend left] \arrow[lu, bend right] \\
%                                               & B                      &                                               
%\end{tikzcd}
%\]
%For $X \in \mathbb{C}$ an $id$-cospan is $X \xrightarrow{id_X} X \xleftarrow{id_X} X$.
Composition of cospans $X \xrightarrow{f} A \xleftarrow{g} Y$ and $Y \xrightarrow{h} B \xleftarrow{i} Z$ is given by pushout: $X \xrightarrow{f} A +_{g,h} B \xleftarrow{i} Z$,  with identity given by the cospan of identities.  Its monoidal product is given by the coproduct of $\mathbb{C}$ as follows: 
\begin{align*}
    (A \xrightarrow{f} \mathcal{G} \xleftarrow{g} B) &\otimes (A' \xrightarrow{f'} \mathcal{G'} \xleftarrow{g'} B') \coloneq\\ A + A' \xrightarrow{f+f'} \mathcal{G} &+ \mathcal{G'} \xleftarrow{g+g'} B + B' ~ , 
\end{align*}
and its unit by the initial object of $\mathbb{C}$.
Symmetry is inherited from $\mathbb{C}$ as the coproduct~\cite{MonoidalCoproduct}.
\end{definition}

Following~\cite{Frobenius2} morphisms of $\textsf{PROP}(\Sigma)$ can be interpreted as particular cospans of discrete hypergraphs,  which we call hypergraphs with interfaces. 
\begin{definition}[Hypergraphs with interfaces]
\label{def:cspd}
Define the category of \emph{hypergraphs with interfaces} $\catname{HypI(\Sigma)}$ as the full sub-category of $\catname{Csp(Hyp(\Sigma))}$ with discrete hypergraphs as objects.
\end{definition}

In particular, this means that any morphism in $\catname{HypI(\Sigma)}$ is of the form $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$,
where $n,\;m$ are discrete hypergraphs and $\mathcal{G}$ is an arbitrary hypergraph.  The morphisms $f$ and $g$ into $\mathcal{G}$ will specify which vertices are to be considered as inputs and outputs,  respectively.   
CHRIS: NOTE THIS IS OK FOR FROBENIUS,  NOT FOR SMC.

@@

In order to represent morphisms of $\textsf{PROP}(\Sigma)$,  it suffices to restrict our attention to MDA-cospans,  as defined below.
\begin{definition}[Category of MDA Hypergraphs with Interfaces \cite{Frobenius2}]
\label{def:monogamy_hyp}
We call a cospan $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ in $\HypI{\Sigma}$ \emph{monogamous directed acyclic (MDA)} if:
\begin{enumerate}
    \item $\mathcal{G}$ is a directed acyclic hypergraph;
    \item $f$ and $g$ are monomorphisms;
    \item the in-degree and out-degree of every vertex is at most $1$;
    \item vertices with in-degree $0$ are precisely the image of $f$; and
    \item vertices with out-degree $0$ are precisely the image of $g$.
\end{enumerate}
We denote by $\MdaCospans$ the subcategory of $\HypI{\Sigma}$ of MDA-cospans. 
\end{definition}
Note that $\MdaCospans$ remains a symmetric monoidal category with the same equipment as before.  In fact,  due to 
\cite{Frobenius2},  we have the following theorem which says the category $\MdaCospans$ in fact \textit{characterizes} $\textsf{PROP}(\Sigma)$. 
\begin{theorem}[Corollary 26~\cite{Frobenius2}]
    $\textsf{PROP}(\Sigma) \cong \MdaCospans$
\end{theorem}
We continue to show a similar theorem relating our \textit{enriched} category $\textsf{PROP}^+(\Sigma)$ to a combinatorial category of e-hypergraphs with interfaces,  introduced in the next section.

CHRIS: EXPLAIN ABOUT MONOIDAL THEORIES AND EQUATIONS.

\subsection{DPOI-Rewriting for Hypergraphs with Interfaces}
CHRIS: PUT BELOW WITH OTHER PRELIMINARIES???

In this section we recall standard results of double-pushout (DPO) rewriting of hypergraphs. 
DPO rewriting is a standard technique for formalising graph rewriting categorically.

@@

To mimic rewrites in $\textsf{PROP}(\Sigma)$ with the help of DPOI rewriting in $\catname{Hyp_{\Sigma}}$, we need to put some restrictions both on pushout complement and on matching as not all of them correspond to valid rewrites in $\textsf{PROP}(\Sigma)$. CHRIS: "LEAD TO"? BE PRECISE! :)

@@

CHRIS: EXPLAIN THE CONDITIONS,  HOW WE WILL INTERPRET THE STRING DIAGRAMS.

We follow analogous steps to those used to achieve a combinatorial representation of $\textsf{PROP}(\Sigma)$ as a particular subset of plain hypergraphs with interfaces.  First we define an e-hypergraph,  which supports a hierarchical notion of "e-box" with distinct components.  Then,  we extend this notion with interfaces via a generalization of the cospan construction.  Finally,  we restrict the morphisms under consideration to those satisfying a certain directed acyclic \textit{monogamy} condition.

>>>>

\begin{figure*}\label{fig:e-graph-example}
\[
	\hspace{1.3cm}
    \scalebox{0.65}{
    \tikzfig{categorical-semantics/egraph-translation}
    }
\]
\caption{Example translation of acyclic e-graphs into string diagrams for semilattice enriched Cartesian categories. }
\end{figure*}

@@

\begin{figure}
\[
	\scalebox{0.65}{
	\tikzfig{categorical-semantics/egraph-strings}
	}
\]
\caption{String diagrams for semilattice enriched symmetric monoidal categories.}
\end{figure}\label{fig:egraph-strings}

@@

Our use of string diagrams is also central to the intuition behind our work.  There is a breadth of work on similar string diagrammatic formalisms,  including hierarchical string diagrams \cite{ghica_hierarchical_2023} and functorial boxes \cite{mellies_functorial_2006},  proof nets for compact closed categories with biproducts \cite{duncan_generalised_2009}, and recent work on tape diagrams for rig categories with biproducts \cite{bonchi_tape_nodate}.

CHRIS: CITE HASEGAWA - TRACE MONOIDAL CATEGORIES, AND PERDIX. 
DAN: I SUGGEST TO CITE HASEGAWA IN FURTHER WORK, FOR GENERALISING THE THEORY TO TRACED MC.
DAN: THIS NEEDS TO BE CAREFULLY EXPANDED TO NOT OFFEND POTENTIAL REVIEWERS.

\section{Categorical Semantics of E-Graphs}
In this section,  we introduce the required preliminaries on semilattice enriched symmetric monoidal categories generated by monoidal theories,  and the string diagrammatic formalism we will use to represent them.

@@

A categorical presentation of a monoidal signature is given by a freely generated products and permutations category.  
\begin{definition}[Products and permutations category]
A \textit{products and permutations category (PROP)} is a strict symmetric monoidal category where every object is the n-fold tensor product of some distinguished object $A$. In this case, we will denote the n-fold tensor product of the distinguished object by simply $n$, so that $n \otimes m = n+m$. A \textit{PROP functor} is a strict SMC-functor which is additionally identity-on-objects. 
We denote the free \textit{products and permutations category} over a monoidal signature $\Sigma$ by $\textbf{PROP}(\Sigma)$.  
\end{definition}
The free category $\textbf{PROP}(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $f \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$, quotiented by the axioms of a symmetric monoidal category.

An equation associated with a monoidal signature is a pair of parallel morphisms of the form above, leading to the following definition of a symmetric monoidal theory. 
\begin{definition}[Symmetric Monoidal Theory]
A \textit{presentation of a symmetric monoidal theory} $(\Sigma, \mathcal{E})$ is given by a pair of a monoidal signature $\Sigma$ and a set $\mathcal{E}$ of pairs of parallel morphisms $\phi,\psi: n \to m$ of $\textbf{PROP}(\Sigma)$.
A \textit{symmetric monoidal theory} $\textbf{SMT}(\Sigma,\mathcal{E})$ generated by a presentation $(\Sigma, \mathcal{E})$ is given by $\textbf{PROP}(\Sigma)$ quotiented by the least congruence including $\mathcal{E}$.
\end{definition}

It will often be convenient to give presentations of SMTs in terms of string diagrams.
\begin{example}
The SMT of \textit{commutative comonoids} is given by the following generators, ${\Delta, !}$, depicted in string diagrammatic form:
\[
	\scalebox{0.8}{
  	 \tikzfig{categorical-semantics/Cartesian-equipment}
	}
\]
together with the following associativity, commutativity and unitality equations, again given in terms of string diagrams. 
\[
	\scalebox{0.6}{
	\tikzfig{categorical-semantics/Cartesian-theory}	
	}
\]
The Cartesian SMT is given by a set of generators $\Sigma_C$, together with the generators and equations of the SMT of commutative monoids, and the following additional naturality equations, for every $c \in \Sigma_C$
\[
	\scalebox{0.6}{
	\tikzfig{categorical-semantics/Cartesian-naturality}
	}
\]
By Fox's Theorem~\cite{fox},  $\textbf{SMT}(\Sigma_C, \mathcal{E}_C)$ is equivalent to the free Cartesian category over $\Sigma_C$. 
\end{example}

@@

We will use enrichment to axiomatize the e-box structure as the ``join" of two morphisms $f,g: A \to B$ as $f + g: A \to B$.\footnote{Note, in particular,  that semilattices do not require a unit for $+$. This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must be a categorical biproduct \cite{maclane} ---a degeneracy we wish to avoid.} More precisely our hom-sets will be given the structure of a semilattice.  
Generally,  we may define a category \textit{enriched} in any other monoidal category $M$ (\textit{e.g.},  \textbf{SLatt}).  In this case,  hom-sets are generalized to hom-objects of $M$ (\textit{e.g.},  a semilattice) and composition 
\[
	\circ: Hom(B,C) \otimes Hom(A,B) \to Hom(A,C)
\]
is defined as a $M$-morphism.  In other words,  composition must respect the additional structure on hom-sets.  
In our case,  the category we wish to enrich is also monoidal,  so we additionally ask that the monoidal structure also respects this structure.  Technically, this amounts to asking that the tensor product is an enriched functor. 
We omit further details of the general case of enrichment here, and instead work with the following more concrete defintion, where $M = \textbf{SLatt}$.

\begin{definition}[Semilattice-enriched PROP]
A \textit{semilattice enriched strict SMC (PROP)}  $\mathbb{C}$ is given by an SMC (PROP) $(\mathbb{C}, \otimes, 1,+)$ where every hom-set additionally has the structure of a semilattice $(\mathbb{C}(A,B), +)$ which respects the composition and tensor product in the following ways,  for all appropriately typed morphisms. 
\begin{align*}
f ; (g+h) &= f;g + f;h &
(f+g) ; h &= f;h + g;h \\
f \otimes (g+h) &= f \otimes g + f \otimes h & 
(f+g) \otimes h &= f \otimes h + g \otimes h,
%\item A set $ob(\mathbb{C})$ of objects;
%\item For every pair $A,B \in ob(\mathbb{C})$, a hom-semigroup $Hom(A,B)$;
%\item An element $\textsf{id}_{A,B} \in Hom(A,B)$; 
\end{align*}
Note,  we take $(;\!)$ and $\otimes$ to bind more tightly than $+$.
A \textit{semilattice enriched strict SMC (PROP) functor} $F: (\mathbb{C}, \otimes, 1,+) \to (\mathbb{D}, \otimes, 1,+)$ is given by a strict SMC (PROP) functor on the underlying SMCs (PROPs) which additionally satisfies $F(f+g) = F(f)+F(g)$.  We denote the \textit{freely enriched PROP (SMT)} over a monoidal signature $\Sigma$ as $\textbf{PROP}^+(\Sigma)$ ($\textbf{SMT}^+(\Sigma, \mathcal{E})$,  respectively). 
\end{definition}

The free category $\textbf{PROP}^+(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $f \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$, as well as $f+g$, quotiented by the axioms of an enriched symmetric monoidal category.

@@

In this section,  we recall the fundamental definitions and results of string diagram rewrite theory for symmetric monoidal categories,  building to the recapitulation of the correspondence between string diagram rewriting and an appropriate notion of double pushout (DPO) rewriting of certain hypergraphs,  as established in~\cite{Frobenius, Frobenius2}.

@@

In ~\cite{Frobenius},  hypergraphs \textit{with interfaces} and their DPO rewriting are shown to be sound and complete for SMTs including a Frobenius algebra. 
Hypergraphs generalize graphs by allowing edges to have multiple sources and targets. 
When interpreting string diagrams as hypergraphs,  morphisms are interpreted as edges,  and wires as vertices.  
However,  there is also a need to specify which vertices are to be considered as \textit{input} and \textit{outputs},  corresponding to the dangling wires at the bottom and top of a string diagram. 
This gives rise to the definition of a hypergraph with interfaces.  
Subsequent work \cite{Frobenius2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous directed acyclic}),  leading to a correspondence between symmetric monoidal string diagrams and the category of such hypergraphs.

\begin{definition}[Category of hypergraphs]\label{def:hypergraph}
A \emph{hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l)$,  where $V$ is a finite set of nodes, $E$ is a finite set of edges, $s : E \to V^{*}$ is a source function, $t : E \to V^{*}$ is a target function,  and $l : E \to \Sigma$ is a labelling function that assigns each edge a generator from monoidal signature $\Sigma$.  The labelling function must respect the typing of the generator: for all edges $e$,  $|s(e)| = m$ and $|t(e)| = n$,  where $l(e) : m \to n$. 
A \emph{hypergraph homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ is given by a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that the following hold. 
\begin{enumerate}
    \item $(s_{\mathcal{F}}\, ;\phi_V^*)(e) = (\phi_E\, ;s_{\mathcal{G}})(e)$
    \item $(t_{\mathcal{F}}\, ;\phi_V^*)(e) = (\phi_E\, ;t_{\mathcal{G}})(e)$
    \item $l_{\mathcal{F}}(e) = \phi_E;l_{\mathcal{G}}(e)$
\end{enumerate}
We denote by $\catname{Hyp(\Sigma)}$ the category of hypergraphs over $\Sigma$.
\end{definition}
Note that $\catname{Hyp(\Sigma)}$ has all finite colimits and, in particular,  the coproduct is given by the disjoint union of hypergraphs.
The initial object is given by the empty hypergraph.

\begin{definition}[Symmetric monoidal category of cospans]
Let $\mathbb{C}$ be a category with all finite colimits.  A cospan from $X$ to $Y$ is a pair of arrows $X \xrightarrow{} A \xleftarrow{} Y$  in $\mathbb{C}$.  This \textit{category of cospans of $\mathbb{C}$},  denoted $\catname{Csp(\mathbb{C})}$,  has as objects those of $\mathbb{C}$ and as arrows isomorphism classes of cospans.
%Two cospans $X \xrightarrow{} A \xleftarrow{} Y$ and $X \xrightarrow{} B \xleftarrow{} Y$ are isomorphic if the following diagram commutes and $\alpha$ is iso.
%\[
%\begin{tikzcd}
%                                               & A \arrow[dd, "\alpha"] &                                                \\
%X \arrow[ru, bend left] \arrow[rd, bend right] &                        & Y \arrow[ld, bend left] \arrow[lu, bend right] \\
%                                               & B                      &                                               
%\end{tikzcd}
%\]
%For $X \in \mathbb{C}$ an $id$-cospan is $X \xrightarrow{id_X} X \xleftarrow{id_X} X$.
Composition of cospans $X \xrightarrow{f} A \xleftarrow{g} Y$ and $Y \xrightarrow{h} B \xleftarrow{i} Z$ is given by pushout: $X \xrightarrow{f} A +_{g,h} B \xleftarrow{i} Z$,  with identity given by the cospan of identities.  Its monoidal product is given by the coproduct of $\mathbb{C}$ as follows: 
\begin{align*}
    (A \xrightarrow{f} \mathcal{G} \xleftarrow{g} B) &\otimes (A' \xrightarrow{f'} \mathcal{G'} \xleftarrow{g'} B') \coloneq\\ A + A' \xrightarrow{f+f'} \mathcal{G} &+ \mathcal{G'} \xleftarrow{g+g'} B + B' ~ , 
\end{align*}
and its unit by the initial object of $\mathbb{C}$.
Symmetry is inherited from $\mathbb{C}$ as the coproduct~\cite{MonoidalCoproduct}.
\end{definition}

Following~\cite{Frobenius2} morphisms of $\textbf{PROP}(\Sigma)$ can be interpreted as particular cospans of discrete hypergraphs,  which we call hypergraphs with interfaces. 
\begin{definition}[Hypergraphs with interfaces]
\label{def:cspd}
Define the category of \emph{hypergraphs with interfaces} $\catname{HypI(\Sigma)}$ as the full sub-category of $\catname{Csp(Hyp(\Sigma))}$ with discrete hypergraphs as objects.
\end{definition}

In particular, this means that any morphism in $\catname{HypI(\Sigma)}$ is of the form $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$,
where $n,\;m$ are discrete hypergraphs and $\mathcal{G}$ is an arbitrary hypergraph.  The morphisms $f$ and $g$ into $\mathcal{G}$ will specify which vertices are to be considered as inputs and outputs,  respectively.

@@

In order to represent only morphisms of $\textbf{PROP}(\Sigma)$,  it is necessary to restrict our attention to the following subset of hypergraphs with interfaces. 
\begin{definition}[Category of MDA Hypergraphs with Interfaces \cite{Frobenius2}]
\label{def:monogamy_hyp}
We call a cospan $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ in $\HypI{\Sigma}$ \emph{monogamous directed acyclic (MDA)} if:
\begin{enumerate}
    \item $\mathcal{G}$ is a directed acyclic hypergraph;
    \item $f$ and $g$ are monomorphisms;
    \item the in-degree and out-degree of every vertex is at most $1$;
    \item vertices with in-degree $0$ are precisely the image of $f$; and
    \item vertices with out-degree $0$ are precisely the image of $g$.
\end{enumerate}
We denote by $\MdaCospans$ the subcategory of $\HypI{\Sigma}$ of MDA-cospans. 
\end{definition}
Note that $\MdaCospans$ remains a symmetric monoidal category with the same equipment as before.  In fact,  due to 
\cite{Frobenius2},  we have the following theorem which says the category $\MdaCospans$ is in fact \textit{equivalent} to $\textbf{PROP}(\Sigma)$. 
\begin{theorem}[Corollary 26~\cite{Frobenius2}]
    $\textbf{PROP}(\Sigma) \cong \MdaCospans$
\end{theorem}

\subsection{DPOI-Rewriting for Hypergraphs with Interfaces}

The correspondence of the previous section further extends to include the equations of a monoidal theory: the equations of a theory are considered as rewrite rules on string diagrams,  which are then modelled as DPO-rewrites on hypergraphs with interfaces. 
DPO rewriting is a standard technique for formalising graph rewriting categorically. 
In this section we recall standard results of double-pushout (DPO) rewriting of hypergraphs,  as well as the specific notion of DPO rewriting which is sound and complete for arbitrary SMTs,  which is called \textit{convex} rewriting.

@@

To mimic rewrites in $\textbf{PROP}(\Sigma)$ with the help of DPOI rewriting in $\catname{Hyp_{\Sigma}}$, we need to put some restrictions both on pushout complement and on matching as not all of them correspond to valid rewrites in $\textbf{PROP}(\Sigma)$. CHRIS: "LEAD TO"? BE PRECISE! :)

@@

CHRIS: EXPLAIN THE CONDITIONS,  HOW WE WILL INTERPRET THE STRING DIAGRAMS.  
REMARK THAT WE CANNOT DEAL WITH SIGNATURES WITH $f:0 \to 0$ morphisms!

We follow analogous steps to those used to achieve a combinatorial representation of $\textbf{PROP}(\Sigma)$ as a particular subset of plain hypergraphs with interfaces.  First we define an e-hypergraph,  which supports a hierarchical notion of "e-box" with distinct components.  Then,  we extend this notion with interfaces via a generalization of the cospan construction.  Finally,  we restrict the morphisms under consideration to those satisfying a certain directed acyclic \textit{monogamy} condition.
